---
title: "Diamond sizes"
date: 2016-08-25
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

---
27.3.1 Exercise 3
---
---
### Something I find striking is that the carat of 2000 diamonds is about 0.3 carats,
### and a little over 2000 diamonds have the carat of 1. 
---

---
27.4.7 Exercise 1
---
---
```{r}

  knitr::opts_chunk$set(
    echo = TRUE
  )
  # Weight of a diamond varied by the cut
  diamonds %>% 
  ggplot(aes(x = carat)) + geom_histogram(bins = 30) + 
  facet_wrap(~cut) + labs(title = "Distribution of Diamonds' Carat By Cut" , x = "Carat" , y = "Number of Diamonds")
  
  # Weight of a diamond varied by clarity
  diamonds %>% ggplot(aes(x = carat)) + 
  geom_histogram(bins = 30) + facet_wrap(~clarity) + labs(title = "Distribution of Diamonds' Carat By Clarity" , x = "Carat" , y = "Number of Diamonds")
  
  # Weight of a diamonds varied by color
  diamonds %>% ggplot(aes(x = carat)) + 
  geom_histogram(bins = 30) + facet_wrap(~color) + labs(title = "Distribution of Diamonds' Carat By Color" , x = "Carat" , y = "Number of Diamonds")
  
```
---
The data concludes that the distribution for 
all carats are left-skewed with the amount of 
the most impressive cuts reaching low weights. 

We can also see that the distribution of carats by clarity also varies, with
those with a clarity of "VS2" having the highest number of small diamonds.

Like the other characteristics, there are a lot of smaller diamonds within this data, but by color, 
they seem slightly more normal than the others. However, there are more diamonds of the color G than other colors.
