---
title: "Predicting Daily Citibike Trips"
author: "Preston Blake"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
library(scales)
```

## Load data

```{r}
trips <- read_tsv('week3/trips_per_day.tsv')

```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
train_set <- trips %>% slice_sample(prop = .8)
vaild_set <- trips %>% slice_sample(prop = .1)
train_n_test_set <- rbind(train_set, vaild_set)
test_set <- anti_join(trips, train_n_test_set, by = "ymd")
```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data

temp_model <- lm(num_trips ~ poly(tmin, degree = 1, raw = T), data = train_set)
temp_train_err <- sqrt(mean((predict(temp_model, train_set) - train_set$num_trips)^2))
temp_valid_err <- sqrt(mean((predict(temp_model, vaild_set) - vaild_set$num_trips)^2))

K <- 1:10
train_err_v <- c()
vaild_err_v <- c()
for (k in K) {
    temp_model <- lm(num_trips ~ poly(tmin, k, raw = T), data = train_set)
    
    # evaluate on the training data
    train_err_v[k] <- sqrt(mean((predict(temp_model, train_set) - train_set$num_trips)^2))

    # evaluate on the validate data
    vaild_err_v[k] <- sqrt(mean((predict(temp_model, vaild_set) - vaild_set$num_trips)^2))
    print(paste0("Training Error for Degree ",k, " is ", train_err_v[k]))
    print(paste0("Validating Error for Degree ",k, " is ", vaild_err_v[k]))
}
```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}

K <- 1:10
train_err_v <- c()
vaild_err_v <- c()
for (k in K) {
    model1 <- lm(num_trips ~ poly(tmin, snow, tmax, degree = k), data = train_set)
    
    # evaluate on the training data
    train_err_v[k] <- sqrt(mean((predict(model1, train_set) - train_set$num_trips)^2))

    # evaluate on the validate data
    vaild_err_v[k] <- sqrt(mean((predict(model1, vaild_set) - vaild_set$num_trips)^2))
    print(paste0("Training Error for Degree ",k, " is ", train_err_v[k]))
    print(paste0("Validating Error for Degree ",k, " is ", vaild_err_v[k]))
}


K <- 1:10
train_err_v <- c()
vaild_err_v <- c()
for (k in K) {
    model1 <- lm(num_trips ~ poly(tmin, snow, prcp, degree = k), data = train_set)
    
    # evaluate on the training data
    train_err_v[k] <- sqrt(mean((predict(model1, train_set) - train_set$num_trips)^2))

    # evaluate on the validate data
    vaild_err_v[k] <- sqrt(mean((predict(model1, vaild_set) - vaild_set$num_trips)^2))
    print(paste0("Training Error for Degree ",k, " is ", train_err_v[k]))
    print(paste0("Validating Error for Degree ",k, " is ", vaild_err_v[k]))
}


K <- 1:10
train_err_v <- c()
vaild_err_v <- c()
for (k in K) {
    model1 <- lm(num_trips ~ poly(tmin, snow, prcp, degree = k), data = train_set)
    
    # evaluate on the training data
    train_err_v[k] <- sqrt(mean((predict(model1, train_set) - train_set$num_trips)^2))

    # evaluate on the validate data
    vaild_err_v[k] <- sqrt(mean((predict(model1, vaild_set) - vaild_set$num_trips)^2))
    print(paste0("Training Error for Degree ",k, " is ", train_err_v[k]))
    print(paste0("Validating Error for Degree ",k, " is ", vaild_err_v[k]))
}

K <- 1:10
train_err_v <- c()
vaild_err_v <- c()
for (k in K) {
    model1 <- lm(num_trips ~ poly(tmin, prcp, snwd, degree = k), data = train_set)
    
    # evaluate on the training data
    train_err_v[k] <- sqrt(mean((predict(model1, train_set) - train_set$num_trips)^2))

    # evaluate on the validate data
    vaild_err_v[k] <- sqrt(mean((predict(model1, vaild_set) - vaild_set$num_trips)^2))
    print(paste0("Training Error for Degree ",k, " is ", train_err_v[k]))
    print(paste0("Validating Error for Degree ",k, " is ", vaild_err_v[k]))
}
```

## Model comparison

Try different models and compare their validation performance.

```{r}
    model1 <- lm(num_trips ~ poly(tmin, snow, tmax, degree = 6), data = train_set)
    # "Training Error for Degree 6 is 4530.49971458804"
    # "Validating Error for Degree 6 is 4614.53482397946"
    # There is a little bit of overfitting here

    model2 <- lm(num_trips ~ poly(tmin, snow, prcp, degree = 4), data = train_set)
    # "Training Error for Degree 4 is 4107.00529687467"
    # "Validating Error for Degree 4 is 4047.48171182731"
    # Out of all the models here, this is the best one due to the errors low gap between
    # and there are no sign of overfitting

    model3 <- lm(num_trips ~ poly(tmin, prcp, snwd, degree = 5), data = train_set)
    # "Training Error for Degree 6 is 3929.14473539237"
    # "Validating Error for Degree 6 is 3878.38460316871"
```

## Final model

Pick your best model and plot the results.

```{r}
model3_RSME <- sqrt(mean((predict(model3, train_set) - train_set$num_trips)^2))


train_set <- train_set %>%
  add_predictions(model3) %>%
  mutate(split = "train")
vaild_set <- vaild_set %>%
  add_predictions(model3) %>%
  mutate(split = "validate")
plot_data <- bind_rows(train_set, vaild_set)

ggplot(plot_data, aes(x = pred, y = num_trips)) +
  geom_point() +
  geom_abline() + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(labels = comma, limits = c(0, 33000)) + 
  labs( x = "Predictions", y = "Actual Number of Trips", title = "Predictions vs Actual Number Of Trips")

# Since there are multiple predictors within this model, I have decided to
# add a predictions vs actual values graph instead.

```

## Save model

```{r}
model3 <- lm(num_trips ~ poly(tmin, prcp, snwd, degree = 4), data = train_set)
```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?


predictions <- predict(model3, test_set)
rmse <- sqrt(mean((predictions - test_set$num_trips)^2))
save(model3, file = 'C:/Users/preston/DS3Coursework/week3/model.Rdata')
# The RSME is around 2.5 million for one sample. However, relative to the data, the RMSEs of the vaildation and training 
# are close, so I don't have a definite prediction on how well it will predict new data.
```
